---
title: "Investigating Exponential random variables"
author: "Jimi Damon"
date: "11/18/2015"
output: pdf_document
hitheme: tomorrow
mode: selfcontained
highlighter: highlight.js
framework: io2012
url:
  assets: ../../assets
  lib: ../../librariesNew
widgets: mathjax
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='./figure/',
                     warning=FALSE, message=FALSE)
```

## Overview

```{r}
set.seed(314159)
lambda=0.2
```

## Simulations 

I make use of the sapply function to return average of a 40 element
exponential function for every element in a list of length 1000.

```{r,echo=TRUE}
sims <- sapply(1:1000,function(x){ mean(rexp(40,lambda)) } )
```

This creates a 1000 element list in the variable sims.

### Sample mean versus Theoretical mean

```{r,echo=TRUE,eval=TRUE}
mean_sims <- mean(sims)
mean_sims
```

The expected mean for an exponential distribution would be $\frac{1}{\lambda}$ which equals

$$\begin{eqnarray*}
E[X] & = & \frac{1}{\lambda} & \newline
     & = & \frac{1}{0.2} & \newline
E[X] & = & 5
\end{eqnarray*}
$$

#### t-Test of this mean value

In order to determine how close this sample mean $\mu$ = 
```r round(mean(sims),2)``` 
is to our ideal value of 5, I will run a t-test.  I will
be testing the null hypothesis that the average of these 1000 
averages of sets of 40 exponential random variables is equivalent to
5.  

I set up the t-test as 

$H_0 :  \mu= 5$

P value is 


### Sample variance vs. Theoretical Variance
```{r,echo=TRUE,eval=TRUE}
var_sims <- var(sims)
var_sims
```
The Expected variance for an exponential random variable would be 
$$
\begin{equation}
\frac{1}{\lambda^2}
\end{equation}
$$

We would get $\frac{1}{(0.2)^2}$ = 25.


### Distribution


3. Distribution is close to normal

```{r}
qplot(sims) + geom_vline(xintercept=mean(sims))
```

The skewness of this distribution tends to favor a fatter tail going to
positive $\infty$. This makes sense since the exponential distribution is limited 
at x = 0  and cannot contribute anymore mass in range $x : \left( -\infty, 0 \right]$.

## Conclusion