---
title: "Investigating Exponential random variables"
author: "Jimi Damon"
date: "11/18/2015"
output: pdf_document
hitheme: tomorrow
mode: selfcontained
highlighter: highlight.js
framework: io2012
url:
  assets: ../../assets
  lib: ../../librariesNew
widgets: mathjax
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='./figure/',
                     warning=FALSE, message=FALSE)
```

## Overview

```{r}
set.seed(314159)
lambda=0.2
```

## Simulations 
```{r,echo=TRUE}
sims <- sapply(1:1000,function(x){ mean(rexp(40,lambda)) } )
```
### Sample mean versus Theoretical mean

```{r,echo=TRUE,eval=TRUE}
mean_sims <- mean(sims)
mean_sims
```

The expected mean for an exponential distribution would be $\frac{1}{\lambda}$ which equals

$$\begin{eqnarray*}
E[X] & = & \frac{1}{\lambda} & \newline
     & = & \frac{1}{0.2} & \newline
E[X] & = & 5
\end{eqnarray*}
$$



### Sample variance vs. Theoretical Variance
```{r,echo=TRUE,eval=TRUE}
var_sims <- var(sims)
var_sims
```
Expected variance would be $\frac{1}{\lambda^2}$



### Distribution


3. Distribution is close to normal

```{r}
hist(sims)
```

The skewness of this distribution tends to favor a fatter tail going ot 
positive $\infty$. This makes sense since the exponential distribution is limited 
at x = 0  and cannot contribute anymore mass in range as $x \to -\infty$.



## Conclusion